---
title: "Data Visualization by Violet Yao"
author: "PS239T"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    df_print: paged
    theme: flatly
    highlight: pygments
---



###  Setup environment

```{r}
# remove all objects
rm(list=ls())

# check working directory
getwd()

# Unload all packages 
library(pacman)
pacman::p_unload(all)

# Add packages 
pacman::p_load(
  tidyverse, #ggplot, dplyr, readr, etc.
  ggthemes, #extra themes
  viridis #great colors
)
```

### Get cleaned dataset
final.RData is the final output by data cleaning rmd.

```{r}
dat <- get(load("../data/final.RData"))
```


```{r}
# Change default plot dimensions
knitr::opts_chunk$set(fig.width=12, fig.height=8)

# Load themes
source("../Set-up/visualisation.R")
```

### Take a look at data
check that it does have character, logical, numeric, integer, and factor variables. 
```{r}
glimpse(dat)
```
```{r}
summary(dat)
```
## ggplot2

## univariate

### univariate 1 (char) - committee_id
Explanation: committee_id is qualitative and categorical, thus I am using bar plot.
```{r}
ggplot(data=dat, aes(committee_id)) + geom_bar( color = "black", fill = "gray")  + 
                         ggtitle("committee_id count")
```

### univariate 2 (logical) - is_house
Explanation: is_house is qualitative and categorical, thus I choose bar plot
```{r}
ggplot(data=dat, aes(is_house)) + geom_bar( color = "black", fill = "gray") +
                         ggtitle("is_house count")
```

### univariate 3 (numeric) - age
Explanation: age is quantitive and continuous, I think histogram, density, and boxplot are all nice representations. My least fav is boxplot because it contains less information, just having 25/50/70 percentile instead of more detailed distribution. 
```{r}
ggplot(data=dat, aes(age)) + geom_histogram(bins = 30, color = "black", fill = "gray") + 
                         ggtitle("age count")

ggplot(data=dat, aes(age)) + geom_density( color = "black", fill = "gray") + 
                         ggtitle("age count")

ggplot(dat, aes(x=factor(1), y=age)) +
  geom_boxplot(width = 0.4, fill = "white") +
  scale_colour_viridis_d(option="A") + 
  labs(x = NULL)  + 
                         ggtitle("age count")
```

### univariate 4 (integer) - birth_year
Explanation: Histogram because birth_year is continuous. 
```{r}
ggplot(data=dat, aes(birth_year)) + geom_histogram(color = "black", fill = "gray") + 
                         ggtitle("birth_year distribution")
```


### univariate 5 (factor) - committee_type
Explanation: Barplot because committee_type is categorical.
```{r}
ggplot(data=dat, aes(committee_type)) + geom_bar( color = "black", fill = "gray")+ 
                         ggtitle("committee_type distribution")
```


## bivariate 

### bivariate 1.1 - age & committee_type - overlayyed histogram
Explanation: Histogram appears most intuitive to me because it contains lots of information - count age distribution, while showing committee_type percentages of each specific age. 
```{r}

ggplot(data=dat) + geom_histogram(aes(x=age, col = committee_type, fill = committee_type), binwidth = 1/4) +
  scale_colour_viridis_d(option="D") +
  ggtitle("age & committee_type")

ggplot(data=dat, aes(age)) + geom_histogram(aes(color = committee_type, fill = committee_type),
                         alpha = 0.4, position = "identity") + 
                         scale_colour_viridis_d(option="A") + 
                         ggtitle("age & committee_type")
```
### bivariate 1.2 - committee_type & age - side-by-side box plot
Explanation: for each category of committee_type, use box plot to show the distribution of age, a continuous variable. 
```{r}
g <- ggplot(dat, aes(committee_type, age))
g + geom_boxplot(varwidth=T, fill="lightgray")  + 
    ggtitle("committee_type & age ")
```

### bivariate 2- committee_id & committee_type
Explanation: when x and y are both categorical data, I am usually thinking of scatterplot. barplot does not appear a good alternative because the height of bar would not mean anything numeric.

```{r}
ggplot(data=dat, aes(x=committee_id, y=committee_type)) + geom_point()
                    
```



## 3 or more variable: committee_name, age, and committee_type
Explanation: Side by side box plot, while boxplot from each committee_type having a different color - provide nice summaries on both age distribution and committee_type for each committee_name.
```{r}
ggplot(data = dat, aes(x = committee_name, y = age, col=committee_type)) +
  geom_boxplot() +
  theme_ed + 
                         ggtitle("commitee_name, age, and committee_type")
```

