---
title: 'Data Cleaning By Violet Yao'
author: "PS239T"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    df_print: paged
    theme: flatly
    highlight: pygments
---

See github page for more information about dataset: https://github.com/unitedstates/congress-legislators/blob





**********************************************************

# SET UP 

```{r}
# remove all objects
rm(list=ls())

# Unload all packages 
library(pacman)
pacman::p_unload(all)

# Add packages 
pacman::p_load(
  tidyverse, #dplyr, readr, etc.
  data.table, #fread() 
  foreign, #load data types including stata .dta files 
  magrittr, #%<>% operator
  skimr #for summerising
)
```

## Check wd

Make sure working directory is datacleaning/code
```{r}
getwd()
```


**********************************************************

# load data

Raw data is in yaml format. I did some parsing in Python and converted them to csv.
```{r}
data_path <- "../data/"

# Read csv of "legislators.csv"
legislators <- read_csv(paste0(data_path, "legislators.csv")) 

# Read csv of  "committee.csv"
committee <- read_csv(paste0(data_path, "committee.csv")) 

# Read csv of "committee_members.csv"
committee_members <- read_csv(paste0(data_path, "committee_members.csv")) 
```

## Save Raw Data
```{r}
save(legislators, committee, committee_members, 
     file="../data/total.RData")
```


**********************************************************
# Look at data 
```{r}
legislators %>% glimpse()

```

```{r}
committee %>% glimpse()
```

```{r}
committee_members %>% glimpse()
```

```{r}
# Rename column for merging
committee %<>% rename(id = thomas_id) 
committee_members %<>% rename(id = comm_id) 
```


## join committee and committee_members
```{r}
# Start with committee as the base dataset and add 
merged <- committee

# Add columns with data from committee, matching on 'id' variable 
merged <- left_join(x = merged,
                          y = committee_members,
                          by = "id")
```


```{r}
# rename column for merging
legislators %<>% rename(leg_id = id) 
```

```{r}
# Add columns with data from legislators, matching on 'leg_id' variable
merged <- left_join(x = merged,
                          y = legislators,
                          by = "leg_id")
```

## check merged
```{r}
merged %>% glimpse()
```


```{r}
# remove dummy column
merged = subset(merged, select=-c(X1.x, X1.y))
```

```{r}
# remove another dummy column
merged = subset(merged, select=-c(X1))
```

```{r}
merged %>% glimpse()
```


## Save combined data
```{r}
save(merged, file="../data/merged.RData")
```

# Summaries

```{r}
merged %>% skimr::skim()
```

## Check missing data. Make sure that your missing data is coded correctly. 
```{r}
# Trim whitespace from ends of all character variables 
merged %<>%
  mutate_if(is.character, list(str_trim))

# Check for empty cells 
merged %>%
  # Keep only character variables 
  select_if(is.character) %>%
  # Recode character variables as 1 if cell is empty 
  mutate_all(list(~ifelse(.=="",1,0))) %>%
  # Add up empty cells for each character variable 
  summarise_all(sum, na.rm=T) %>%
  # Transpose data for visibility 
  t() 
```

## rename confusing columns
```{r}
colnames(merged)[colnames(merged) == 'name'] <- 'committee_name'
colnames(merged)[colnames(merged) == 'type'] <- 'committee_type'
colnames(merged)[colnames(merged) == 'id'] <- 'committee_id'
colnames(merged)[colnames(merged) == 'leg_id'] <- 'legislator_id'
```

## take a look at merged after renaming
```{r}
merged %>% glimpse()
```

# Create five or more variables using mutate().
```{r}
merged %<>%
  # Turn race into factor var
  mutate(committee_type = as.factor(committee_type)) %>% 
  # Create binary variable for suicide
  mutate(is_house = ifelse(committee_type == "house", T, F))
```

```{r}
merged %<>%
  # Add year variable 
  mutate(birth_year = year(birthday)) %>%
  # Add year variable 
  mutate(birth_month = month(birthday)) %>%
    # Add year variable 
  mutate(birth_day = mday(birthday))
```

```{r}
merged %<>%
  # Add year variable 
  mutate(age = 2019 - birth_year) 
```

create new variable legislator_age using base r 
```{r}
merged$legislator_age <- 2019 - merged$birth_year
```

As above, I created more than variables using mutate().

These variables include at least one of each data type (character, logical, numeric, and integer) as well as at least one factor variable. 
For lesgislator_age, I showed that I created with base r instead of dplyr. 
```{r}
merged %>% glimpse()
```


### Generate a summary statistic for at least five variables using summarize().

```{r}
merged %>% glimpse()


merged %>% 
  summarise_if(is.numeric, mean, na.rm=T)

merged %>% 
  group_by(committee_type) %>%
  summarise(avg=mean(age, na.rm=T),
            sd=sd(age, na.rm=T))

merged %>% 
  group_by(committee_name) %>%
  summarise(avg=mean(age, na.rm=T),
            sd=sd(age, na.rm=T))

merged %>% 
  group_by(committee_type, committee_name) %>%
  summarise(n=n()) 

merged %>% 
  group_by(is_house) %>%
  summarise(n=n()) 
```
```{r}
# use base r for summarizing legislator age
summary(merged$legislator_age)
```


## use of spread and gather
```{r}
# Frequency of weapons 
grouped_by_name<- merged %>% 
  group_by(committee_type, committee_name) %>% 
  summarise(total=n()) 


head(grouped_by_name)
# Spread out totals by year 
grouped_by_name %>% 
  spread(key=committee_name, value=total)

head(grouped_by_name)
# Restack data 
grouped_by_name %>% 
  spread(key=committee_name, value=total) %>%
  gather(key=committee_name, value=total, -committee_type)

head(grouped_by_name)
```

save data for visualizations

```{r}
save(merged, file="../data/final.RData")
```






















